---
title: "Surface chamber radiocarbon plots"
author: "Lydia Vaughn"
date: "January 24, 2018"
output: html_document
---

Data: atmosphere_corrected_chamber_radiocarbon.csv is the analyzed dataset produced with the surface_chamber_atm_correction.Rmd code.  loss_corrected_bulk_soil_radiocarbon.csv is the dataset produced with the bulk_soil_14C_plots.Rmd code. flux_CO2_CH4_Barrow_2012_2014.csv includes all associated surface CO2 flux data and can be accessed at http://dx.doi.org/10.5440/1364062
```{r}
atmcorr <- read.csv('data/atmosphere_corrected_chamber_radiocarbon.csv', header = T, sep = ',', stringsAsFactors = F)
headers = read.csv('data/flux_CO2_CH4_Barrow_2012_2014.csv', skip = 5, header = F, nrows = 1, as.is = T)
flux <- read.csv('data/flux_CO2_CH4_Barrow_2012_2014.csv', skip = 7, header = F, sep = ',', stringsAsFactors = F)
colnames(flux) <- headers
bulk <- read.csv('data/loss_corrected_bulk_soil_radiocarbon.csv', header = T, sep = ',', stringsAsFactors = F)
```

```{r}
library(tidyverse)
library(lubridate)
library(gridExtra)
```

get rid of low quality data; remove any measurements for which fs is less than 0.5
```{r}
highf <- atmcorr %>% filter(f_soil >= 0.5, !is.na(position))
```

Design a plot themes
```{r}
boxtheme <- theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  #theme(panel.border = element_rect(color="black", size=0.5, linetype="solid")) +
  theme(axis.text.y = element_text(color="black", size=9)) +
  theme(axis.text.x = element_text(color="black", size=11)) +
  theme(axis.title.y = element_text(size=11)) +
  theme(axis.title.x = element_blank()) 

flux14Ctheme <- theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  #theme(panel.border = element_rect(color="black", size=0.5, linetype="solid")) +
  theme(axis.text = element_text(color="black", size=9)) +
  theme(axis.title = element_text(color="black", size=11)) 

theme2014 <- theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  #theme(panel.border = element_rect(color="black", size=0.5, linetype="solid")) +
  theme(axis.text.y = element_text(color="black", size=9)) +
  theme(axis.text.x = element_text(color="black", size=10)) +
  theme(axis.title.y = element_text(size=11)) +
  theme(axis.title.x = element_blank()) 

```

plot all highf data, separated by position
```{r}
position <- ggplot(highf, aes(x=position, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  scale_x_discrete(breaks=c("Center","Edge","Trough"), labels=c("Center","Rim","Trough")) 

print(position + boxtheme)
ggsave("chamber14C_position.png", path="plots", width = 6,height = 4, units="in", dpi = 300)
```

plot all highf data, separated by type
```{r}
type <- ggplot(highf, aes(x=type, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered")) 


print(type + boxtheme)
ggsave("chamber14C_type.png", path="plots", width = 6,height = 4, units="in", dpi = 300)
```

Facet the above plot by polygon position
```{r}
type <- ggplot(highf, aes(x=type, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  facet_grid(. ~ position)

print(type + boxtheme)
```

Plot September 2014 data, separated by type
```{r}
type1409 <- ggplot(highf %>% filter(obs_year == "2014"), aes(x=type, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered")) 


print(type1409 + boxtheme)
ggsave("chamber14C_type_1409.png", path="plots", width = 6,height = 4, units="in", dpi = 300)
```

Make a bar graph of the above data
```{r}
summ1409 <- highf %>% filter(obs_year == "2014") %>% group_by(type) %>% summarize(meanC14 = mean(C14_atmcorr_permil), sdC14 = sd(C14_atmcorr_permil), nC14 = length(C14_atmcorr_permil)) %>% mutate(seC14 = sdC14/nC14^0.5)

bar2014<-ggplot(summ1409, aes(x=type, y=meanC14, fill = type)) +
  geom_hline(yintercept=17.7, linetype="dashed", size=0.3) +
  geom_bar(stat="identity", width=0.7) +
  geom_errorbar(aes(ymin=meanC14-seC14, ymax=meanC14+seC14), position=position_dodge(), width=0.2, size=0.4) +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  geom_hline(yintercept=0, size=0.3) +
  scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered"))  +
  ggtitle('c') +
  scale_fill_manual(breaks=c("H","F","L"), labels=c("High-centered","Flat-centered","Low-centered"), values = c("darkslateblue", "aquamarine4", "darkgoldenrod1" ), name = "Polygon type") +
  theme(plot.title = element_text(colour = "black", size=14, hjust = -0.1)) 


print(bar2014 + boxtheme) 
```

Plot the CO2 fluxes from September 2014 as a bar chart
```{r}
flux$observation_date <- ymd(flux$observation_date)
flux$month <- format(flux$observation_date, "%Y-%m")

summflux <- flux %>% filter(month == "2014-08" | month == "2014-09") %>% group_by(type) %>% summarize(meanCO2 = mean(flux_CO2), sdCO2 = sd(flux_CO2), nCO2 = length(flux_CO2)) %>% mutate(seCO2 = sdCO2/nCO2^0.5)

flux2014<-ggplot(summflux, aes(x=type, y=meanCO2, fill = type)) +
  geom_bar(stat="identity", width=0.7) +
  geom_errorbar(aes(ymin=meanCO2-seCO2, ymax=meanCO2+seCO2), position=position_dodge(), width=0.2, size=0.4) +
  ylab(expression(Ecosystem~Respiration~(mu*mol~CO[2]~~m^{-2}~~s^{-1}))) +
  geom_hline(yintercept=0, size=0.3) +
    scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered"))  +
  scale_fill_manual(breaks=c("H","F","L"), labels=c("High-centered","Flat-centered","Low-centered"), values = c("darkslateblue", "aquamarine4", "darkgoldenrod1" ), name = "Polygon type") +
    ggtitle('b') 

print(flux2014 +boxtheme) 
```

Plot thaw depth from September 2014 as a bar chart
```{r}
summthaw <- flux %>% filter(month == "2014-08" | month == "2014-09") %>% group_by(type) %>% summarize(mean.thaw = mean(thaw_depth), sd.thaw = sd(thaw_depth), n.thaw = length(thaw_depth)) %>% mutate(se.thaw = sd.thaw/n.thaw^0.5)

thaw2014<-ggplot(summthaw, aes(x=type, y=mean.thaw, fill = type)) +
  geom_bar(stat="identity", width=0.7) +
  geom_errorbar(aes(ymin=mean.thaw-se.thaw, ymax=mean.thaw+se.thaw), position=position_dodge(), width=0.2, size=0.4) +
  #ylab("Thaw depth (cm)") +
  ylab(expression(Thaw~Depth~(cm))) +
  geom_hline(yintercept=0, size=0.3) +
  scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered"))  +
  scale_fill_manual(breaks=c("H","F","L"), labels=c("High-centered","Flat-centered","Low-centered"), values = c("darkslateblue", "aquamarine4", "darkgoldenrod1" ), name = "Polygon type") +
  ggtitle('a') 

print(thaw2014 + boxtheme) 
 
```

Plot bulk soil radiocarbon from September 2014 as a bar chart
```{r}
toplot <- bulk %>% filter(O2 == "aerobic") %>% separate(sample_name, into = c("prof", "ancillary"), sep = "-") %>% separate(ancillary, into = c("depth", "year"), sep = " ") %>% separate(name, into = c("type", "position"), by = "-") %>% separate(type, into = c("type", "plot_number"), sep = "C")

summ <- toplot %>% filter(depth == "1") %>% group_by(type) %>% summarize(meanC14 = mean(C14_total, na.rm=T), sdC14 = sd(C14_total, na.rm=T), nC14 = length(C14_total[!is.na(C14_total)])) %>% mutate(seC14 = sdC14/nC14^0.5) %>% add_column(increment = "shallow") %>% full_join(toplot %>% filter(depth == "2") %>% group_by(type) %>% summarize(meanC14 = mean(C14_total, na.rm=T), sdC14 = sd(C14_total, na.rm=T), nC14 = length(C14_total)) %>% mutate(seC14 = sdC14/nC14^0.5) %>% add_column(increment = "deep"))

summ$seC14 <- ifelse(summ$nC14 > 2, summ$seC14, NA)

barbulk<-ggplot(summ, aes(x=increment, y=meanC14, fill=type)) +
  geom_hline(yintercept=17.7, linetype="dashed", size=0.3) +
  geom_bar(stat="identity", width=0.7, position = position_dodge(0.7)) +
  geom_errorbar(aes(ymin=meanC14-seC14, ymax=meanC14+seC14), position=position_dodge(0.7), width=0.2, size=0.4) +
  ylab(expression(Delta^14*C[soil] ~ ("\u2030"))) +
  geom_hline(yintercept=0, size=0.3) +
  scale_fill_manual(breaks=c("F","H","L"), labels=c("Flat-centered","High-centered","Low-centered"), values = c("darkslateblue", "aquamarine4", "darkgoldenrod1" ), name = "Polygon type") +
  scale_x_discrete(breaks = c("shallow", "deep"), labels = c("Shallow", "Deep")) +
  ggtitle('d') 


print(barbulk + boxtheme + theme(legend.position = c(.75, .25)))

#ggsave("bulksoil14C.png", path="plots", width = 6,height = 5, units="in", dpi = 300)
```

Arrange the plots for publication
```{r}
p1 <- thaw2014 + theme2014 + theme(axis.title.y = element_text(vjust=2)) + theme(plot.margin = unit(c(3, 5, 3, 4), "mm"))
p2 <- flux2014 + theme2014 + theme(plot.margin = unit(c(3, 2, 3, 6), "mm"))
p3 <- bar2014 + theme2014 + theme(plot.margin = unit(c(3, 5, 3, 3), "mm"))
p4 <- barbulk + theme2014 + theme(legend.position = c(.75, .25)) + theme(plot.margin = unit(c(3, 2, 3, 4), "mm"))
combo <- grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2, heights = c(8,8))
#combo <- grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2, heights=c(8,8,8))
ggsave("plots2014.png", combo, path="plots", width = 8,height =7,dpi = 300)
```

Make a dataframe that combines the flux data and radiocarbon data
```{r}
highf$observation_date <- ymd(highf$observation_date)
highf$month <- format(highf$observation_date, "%Y-%m")

flux14C <- highf %>% select(plot_ID,month,C14_atmcorr_permil) %>% inner_join(flux %>% filter(chamber_type == "opaque") %>% mutate(month = ifelse(month == "2014-08", "2014-09", month)) %>%  select(plot_ID,month,flux_CO2))
```

Make a scatter plot of all measurements
```{r}
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

flux14Cplot <- ggplot(flux14C, aes(y=C14_atmcorr_permil, x=flux_CO2, colour=month, shape=month)) +
  geom_point(size=3) +
  ylab(expression(Delta^14*C[Reco]~("\u2030"))) +
  scale_colour_manual(values=cbPalette, name="Month", labels=c("July 2013", "September 2013", "September 2014")) +
  scale_shape_discrete(name="Month", labels=c("July 2013", "September 2013", "September 2014")) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  xlab(expression("Ecosystem respiration"~(mu*mol~CO[2]~m^-2~y^-1))) 

print(flux14Cplot + flux14Ctheme + theme(legend.position = c(.75, .25)) 
)
ggsave("chamber14C_flux.png", path="plots", width = 6,height = 5, units="in", dpi = 300)
```

Add a doy variable to the highf dataframe
```{r}
highf$doy <- yday(highf$observation_date)
```

Plot all highf 14C measurements by month
```{r}
monthplot <- ggplot(highf, aes(y=C14_atmcorr_permil, x=month)) +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco]~("\u2030"))) +
  ggtitle("all surface chamber measurements")
print(monthplot)
```

With the above plot, combine data from different years and separate polygons by type
```{r}
types <- c(
  flat = "Flat-centered",
  high = "High-centered",
  low = "Low-centered"
)

plot_labeller <- function(value){
  value[value=="flat"] <- "Flat-centered"
  value[value=="high"] <- "High-centered"
  value[value=="low"] <- "Low-centered"
  return(value)
}

striptheme <- theme(strip.background = element_rect(colour="white",fill="white")) +
  theme(panel.spacing = unit(0.5,"lines")) +
  theme(strip.text = element_text(size=12))

monthplot_type <- ggplot(highf %>% mutate(month = format(observation_date, "%B"), month = ordered(month, levels = c("July","August","September","October"))), aes(y=C14_atmcorr_permil, x=month)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  facet_grid(.~type, labeller=plot_labeller) +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) 

print(monthplot_type + boxtheme + striptheme + theme(axis.text.x = element_text(angle=90,vjust=0.5)))
ggsave("month_type_14C.png", path="plots", width = 6,height = 4, units="in", dpi = 300)
```

same plot as above but as a scatterplot
```{r}
monthplot_type_scatter <- ggplot(highf %>% mutate(month = format(observation_date, "%B"), month = ordered(month, levels = c("July","August","September","October"))), aes(y=C14_atmcorr_permil, x=month)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_point() +
  facet_grid(.~type, labeller=plot_labeller) +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) 
  

print(monthplot_type_scatter + flux14Ctheme + striptheme + theme(axis.text.x = element_text(angle=90, size=10, color="black", vjust=0.5)))
```

Same plot but with doy on the x axis
```{r}
doyplot_type_scatter <- ggplot(highf %>% mutate(month = format(observation_date, "%B"), month = ordered(month, levels = c("July","August","September","October"))), aes(y=C14_atmcorr_permil, x=doy)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_point() +
  facet_grid(.~type, labeller=plot_labeller) +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) 
  

print(doyplot_type_scatter + flux14Ctheme + striptheme)
ggsave("doy_type_scatter_14C.png", path="plots", width = 6,height = 4, units="in", dpi = 300)
```

Remove faceting
```{r}
monthplot_scatter <- ggplot(highf %>% mutate(month = format(observation_date, "%B"), month = ordered(month, levels = c("July","August","September","October"))), aes(y=C14_atmcorr_permil, x=doy)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) 

print(monthplot_scatter + flux14Ctheme)
```