---
title: "Surfac chamber radiocarbon plots"
author: "Lydia Vaughn"
date: "January 24, 2018"
output: html_document
---

Data: atmosphere-corrected chamber radiocarbon data and surface CO2 flux data (available at http://dx.doi.org/10.5440/1364062)
```{r}
atmcorr <- read.csv('data/atmosphere_corrected_chamber_radiocarbon.csv', header=T, sep=',', stringsAsFactors=F)
flux <- read.csv('data/flux.csv', header=T, sep=',', stringsAsFactors=F)
```

```{r}
library(tidyverse)
library(lubridate)
library(gridExtra)
```

get rid of low quality data; remove any measurements for which fs is less than 0.5
```{r}
highf <- atmcorr %>% filter(f_soil >= 0.5, !is.na(position))
```

Design a theme for all plots
```{r}
boxtheme <- theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  #theme(panel.border = element_rect(color="black", size=0.5, linetype="solid")) +
  theme(axis.text.y = element_text(color="black", size=12)) +
  theme(axis.text.x = element_text(color="black", size=14)) +
  theme(axis.title.y = element_text(size=14)) +
  theme(plot.title = element_text(colour = "black", size=14, hjust = -0.1)) +
  theme(axis.title.x = element_blank()) +
  theme(plot.title = element_text(colour = "black", size=14, hjust = 0)) 


```

plot all high f_soil data, separated by position
```{r}
position <- ggplot(highf, aes(x=position, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  scale_x_discrete(breaks=c("Center","Edge","Trough"), labels=c("Center","Rim","Trough")) 

print(position + boxtheme)
ggsave("chamber14C_position.png", path="plots", width = 6,height = 4, units="in", dpi = 600)
```

plot all high f_soil data, separated by type
```{r}
type <- ggplot(highf, aes(x=type, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered")) 


print(type + boxtheme)
ggsave("chamber14C_type.png", path="plots", width = 6,height = 4, units="in", dpi = 600)
```

Facet the above plot by polygon position
```{r}
type <- ggplot(highf, aes(x=type, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  facet_grid(. ~ position)

print(type + boxtheme)
```

Plot September 2014 data, separated by type
```{r}
type1409 <- ggplot(highf %>% filter(obs_year == "2014"), aes(x=type, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered")) 


print(type1409 + boxtheme)
ggsave("chamber14C_type_1409.png", path="plots", width = 6,height = 4, units="in", dpi = 600)
```

Make a bar graph of the above data
```{r}
summ1409 <- highf %>% filter(obs_year == "2014") %>% group_by(type) %>% summarize(meanC14 = mean(C14_atmcorr_permil), sdC14 = sd(C14_atmcorr_permil), nC14 = length(C14_atmcorr_permil)) %>% mutate(seC14 = sdC14/nC14^0.5)

bar2014<-ggplot(summ1409, aes(x=type, y=meanC14)) +
  geom_hline(yintercept=17.7, linetype="dashed", size=0.3) +
  geom_bar(stat="identity", width=0.7) +
  geom_errorbar(aes(ymin=meanC14-seC14, ymax=meanC14+seC14), position=position_dodge(), width=0.2, size=0.4) +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  geom_hline(yintercept=0, size=0.3) +
    scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered"))  +
    ggtitle('b') 

print(bar2014 +boxtheme) 
```

Plot the CO2 fluxes from September 2014 as a bar chart
```{r}
flux$observation_date <- ymd(flux$observation_date)
flux$month <- format(flux$observation_date, "%Y-%m")

summflux <- flux %>% filter(month == "2014-08" | month == "2014-09") %>% group_by(type) %>% summarize(meanCO2 = mean(flux_CO2), sdCO2 = sd(flux_CO2), nCO2 = length(flux_CO2)) %>% mutate(seCO2 = sdCO2/nCO2^0.5)

flux2014<-ggplot(summflux, aes(x=type, y=meanCO2)) +
  geom_bar(stat="identity", width=0.7) +
  geom_errorbar(aes(ymin=meanCO2-seCO2, ymax=meanCO2+seCO2), position=position_dodge(), width=0.2, size=0.4) +
  ylab(expression(Ecosystem~Respiration~(mu*mol~CO[2]~~m^{-2}~~s^{-1}))) +
  geom_hline(yintercept=0, size=0.3) +
    scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered"))  +
    ggtitle('a') 

print(flux2014 +boxtheme) 
```

Arrange the plots for publication
```{r}
p1 <- flux2014 + boxtheme
p2 <- bar2014 + boxtheme
combo <- grid.arrange(p1, p2, nrow=2, heights=c(8,8))
ggsave("plots2014.png", combo, path="plots", width = 8,height = 11,dpi = 1200)
```
