---
title: "Surface chamber radiocarbon plots"
author: "Lydia Vaughn"
date: "January 24, 2018"
output: html_document
---

Data: atmosphere_corrected_chamber_radiocarbon.csv is the analyzed dataset produced with the surface_chamber_atm_correction.Rmd code. flux_CO2_CH4_Barrow_2012_2014.csv includes all associated surface CO2 flux data and can be accessed at http://dx.doi.org/10.5440/1364062
```{r}
atmcorr <- read.csv('data/atmosphere_corrected_chamber_radiocarbon.csv', header=T, sep=',', stringsAsFactors=F)
flux <- read.csv('data/flux_CO2_CH4_Barrow_2012_2014.csv', header=T, sep=',', stringsAsFactors=F)
```

```{r}
library(tidyverse)
library(lubridate)
library(gridExtra)
```

get rid of low quality data; remove any measurements for which fs is less than 0.5
```{r}
highf <- atmcorr %>% filter(f_soil >= 0.5, !is.na(position))
```

Design a plot themes
```{r}
boxtheme <- theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  #theme(panel.border = element_rect(color="black", size=0.5, linetype="solid")) +
  theme(axis.text.y = element_text(color="black", size=12)) +
  theme(axis.text.x = element_text(color="black", size=14)) +
  theme(axis.title.y = element_text(size=14)) +
  theme(plot.title = element_text(colour = "black", size=14, hjust = -0.1)) +
  theme(axis.title.x = element_blank()) +
  theme(plot.title = element_text(colour = "black", size=14, hjust = 0))

flux14Ctheme <- theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  #theme(panel.border = element_rect(color="black", size=0.5, linetype="solid")) +
  theme(axis.text = element_text(color="black", size=12)) +
  theme(axis.title = element_text(color="black", size=14)) 

```

plot all highf data, separated by position
```{r}
position <- ggplot(highf, aes(x=position, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  scale_x_discrete(breaks=c("Center","Edge","Trough"), labels=c("Center","Rim","Trough")) 

print(position + boxtheme)
ggsave("chamber14C_position.png", path="plots", width = 6,height = 4, units="in", dpi = 600)
```

plot all highf data, separated by type
```{r}
type <- ggplot(highf, aes(x=type, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered")) 


print(type + boxtheme)
ggsave("chamber14C_type.png", path="plots", width = 6,height = 4, units="in", dpi = 600)
```

Facet the above plot by polygon position
```{r}
type <- ggplot(highf, aes(x=type, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  facet_grid(. ~ position)

print(type + boxtheme)
```

Plot September 2014 data, separated by type
```{r}
type1409 <- ggplot(highf %>% filter(obs_year == "2014"), aes(x=type, y=C14_atmcorr_permil)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered")) 


print(type1409 + boxtheme)
ggsave("chamber14C_type_1409.png", path="plots", width = 6,height = 4, units="in", dpi = 600)
```

Make a bar graph of the above data
```{r}
summ1409 <- highf %>% filter(obs_year == "2014") %>% group_by(type) %>% summarize(meanC14 = mean(C14_atmcorr_permil), sdC14 = sd(C14_atmcorr_permil), nC14 = length(C14_atmcorr_permil)) %>% mutate(seC14 = sdC14/nC14^0.5)

bar2014<-ggplot(summ1409, aes(x=type, y=meanC14)) +
  geom_hline(yintercept=17.7, linetype="dashed", size=0.3) +
  geom_bar(stat="identity", width=0.7) +
  geom_errorbar(aes(ymin=meanC14-seC14, ymax=meanC14+seC14), position=position_dodge(), width=0.2, size=0.4) +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) +
  geom_hline(yintercept=0, size=0.3) +
    scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered"))  +
    ggtitle('b') 

print(bar2014 +boxtheme) 
```

Plot the CO2 fluxes from September 2014 as a bar chart
```{r}
flux$observation_date <- ymd(flux$observation_date)
flux$month <- format(flux$observation_date, "%Y-%m")

summflux <- flux %>% filter(month == "2014-08" | month == "2014-09") %>% group_by(type) %>% summarize(meanCO2 = mean(flux_CO2), sdCO2 = sd(flux_CO2), nCO2 = length(flux_CO2)) %>% mutate(seCO2 = sdCO2/nCO2^0.5)

flux2014<-ggplot(summflux, aes(x=type, y=meanCO2)) +
  geom_bar(stat="identity", width=0.7) +
  geom_errorbar(aes(ymin=meanCO2-seCO2, ymax=meanCO2+seCO2), position=position_dodge(), width=0.2, size=0.4) +
  ylab(expression(Ecosystem~Respiration~(mu*mol~CO[2]~~m^{-2}~~s^{-1}))) +
  geom_hline(yintercept=0, size=0.3) +
    scale_x_discrete(breaks=c("Flat","High","Low"), labels=c("Flat-centered","High-centered","Low-centered"))  +
    ggtitle('a') 

print(flux2014 +boxtheme) 
```

Arrange the plots for publication
```{r}
p1 <- flux2014 + boxtheme
p2 <- bar2014 + boxtheme
combo <- grid.arrange(p1, p2, nrow=2, heights=c(8,8))
ggsave("plots2014.png", combo, path="plots", width = 8,height = 11,dpi = 1200)
```

Make a dataframe that combines the flux data and radiocarbon data
```{r}
highf$observation_date <- ymd(highf$observation_date)
highf$month <- format(highf$observation_date, "%Y-%m")

flux14C <- highf %>% select(plot_ID,month,C14_atmcorr_permil) %>% inner_join(flux %>% filter(chamber_type == "opaque") %>% mutate(month = ifelse(month == "2014-08", "2014-09", month)) %>%  select(plot_ID,month,flux_CO2))
```

Make a scatter plot of all measurements
```{r}
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

flux14Cplot <- ggplot(flux14C, aes(y=C14_atmcorr_permil, x=flux_CO2, colour=month, shape=month)) +
  geom_point(size=2) +
  ylab(expression(Delta^14*C[Reco]~("\u2030"))) +
  scale_colour_manual(values=cbPalette, name="Month") +
  scale_shape_discrete(name="Month") +
  xlab(expression(R[eco]~(mu*mol~CO[2]~m^-2~y^-1))) 

print(flux14Cplot + flux14Ctheme)
ggsave("chamber14C_flux.png", path="plots", width = 6,height = 4, units="in", dpi = 600)
```

Add a doy variable to the highf dataframe
```{r}
highf$doy <- yday(highf$observation_date)
```

Plot all highf 14C measurements by month
```{r}
monthplot <- ggplot(highf, aes(y=C14_atmcorr_permil, x=month)) +
  geom_boxplot() +
  geom_point() +
  ylab(expression(Delta^14*C[Reco]~("\u2030"))) +
  ggtitle("all surface chamber measurements")
print(monthplot)
```

With the above plot, combine data from different years and separate polygons by type
```{r}
types <- c(
  flat = "Flat-centered",
  high = "High-centered",
  low = "Low-centered"
)

plot_labeller <- function(value){
  value[value=="flat"] <- "Flat-centered"
  value[value=="high"] <- "High-centered"
  value[value=="low"] <- "Low-centered"
  return(value)
}

striptheme <- theme(strip.background = element_rect(colour="white",fill="white")) +
  theme(panel.spacing = unit(0.5,"lines")) +
  theme(strip.text = element_text(size=14))

monthplot_type <- ggplot(highf %>% mutate(month = format(observation_date, "%B"), month = ordered(month, levels = c("July","August","September","October"))), aes(y=C14_atmcorr_permil, x=month)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_boxplot() +
  geom_point() +
  facet_grid(.~type, labeller=plot_labeller) +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) 
  theme(axis.text.x = element_text(angle=90, size=10, color="black", vjust=0.5)) +

print(monthplot_type + boxtheme + striptheme + theme(axis.text.x = element_text(angle=90,vjust=0.5)))
ggsave("month_type_14C.png", path="plots", width = 6,height = 4, units="in", dpi = 600)
```

Same plot but as a scatterplot with doy on the x axis
```{r}
monthplot_type_scatter <- ggplot(highf %>% mutate(month = format(observation_date, "%B"), month = ordered(month, levels = c("July","August","September","October"))), aes(y=C14_atmcorr_permil, x=doy)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_point() +
  facet_grid(.~type, labeller=plot_labeller) +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) 
  theme(axis.text.x = element_text(angle=90, size=10, color="black", vjust=0.5)) +

print(monthplot_type_scatter + flux14Ctheme + striptheme)
```

Remove faceting
```{r}
monthplot_scatter <- ggplot(highf %>% mutate(month = format(observation_date, "%B"), month = ordered(month, levels = c("July","August","September","October"))), aes(y=C14_atmcorr_permil, x=doy)) +
  geom_hline(yintercept=19.7, size=0.3, linetype="dashed") +
  geom_point() +
  ylab(expression(Delta^14*C[Reco] ~ ("\u2030"))) 
  theme(axis.text.x = element_text(angle=90, size=10, color="black", vjust=0.5)) +

print(monthplot_scatter + flux14Ctheme)
```