---
title: "Soil profile radiocarbon plots"
author: "Lydia Vaughn"
date: "29/01/2018"
output: html_document
---

Data: atmosphere_corrected_chamber_radiocarbon.csv is the analyzed dataset produced with the surface_chamber_atm_correction.Rmd code.  naming.csv assigns a new name to each chamber according to the naming convention for this analysis. radiocarbon_field_Barrow_2012_2013_2014.csv includes all non-corrected radiocarbon data from surface chambers and soil pore-space CO2 samples.  Dataset can be accessed at http://dx.doi.org/10.5440/1364062
```{r}
atmcorr <- read.csv('data/atmosphere_corrected_chamber_radiocarbon.csv', header=T, sep=',', stringsAsFactors=F)
noncorr <- read.csv('data/radiocarbon_field_Barrow_2012_2013_2014.csv', header=T, sep=',', stringsAsFactors=F)
naming <- read.csv('data/naming.csv', header=T, sep=',', stringsAsFactors=F)
```

```{r}
library(tidyverse)
library(lubridate)
```

Add the new chamber names to both dataframes
```{r}
atmcorr <- atmcorr %>% left_join(naming)
noncorr <- noncorr %>% left_join(naming)
```

Make a variable in each data frame for profile and sampling month
```{r}
atmcorr <- atmcorr %>% mutate(observation_date = ymd(observation_date), month = format(observation_date, "%Y-%m"), measurement = paste(name, month)) 

noncorr <- noncorr %>% mutate(observation_date = ymd(observation_date), month = format(observation_date, "%Y-%m"), measurement = paste(name, month)) 
```

With the atmosphere-corrected data, join to the profile data, limiting the join to only samples where profile data are available
```{r}
profiles_atmcorr <- atmcorr %>% filter(!is.na(C14_atmcorr_permil) & measurement %in% noncorr[noncorr$material == "soil pore gas",]$measurement) %>% mutate(X14C = C14_atmcorr_permil, X14C_sigma = C14_atmcorr_permil_sd) %>% full_join(noncorr %>% filter(material == "soil pore gas"))
```

Limit non-atmosphere-corrected data to only profiles that include subsurface samples
```{r}
profiles_noncorr <- noncorr %>% filter(material=="surface emissions" & measurement %in% noncorr[noncorr$material == "soil pore gas",]$measurement) %>% full_join(noncorr %>% filter(material == "soil pore gas"))
```


Design plot theme
```{r}
proftheme <-  theme_bw() +
  theme(panel.border = element_rect(color="black", size=0.5, linetype="solid")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  theme(axis.text = element_text(color="black", size=9)) +
  theme(axis.title = element_text(color="black", size=11)) +
  theme(legend.text = element_text(color="black", size=9)) +
  theme(legend.title = element_text(color="black", size=10))
```


Plot 14C depth profiles using the atmosphere-corrected data
```{r}
profiles_atmcorr <- profiles_atmcorr %>% mutate(depth = ifelse(!is.na(depth), depth, 0)) %>% arrange(measurement, depth)

atmcorrplot <- ggplot(profiles_atmcorr, aes(x=X14C, y=depth, group=measurement)) +
  geom_point(aes(colour=measurement)) +
  geom_path(aes(colour=measurement)) +
  scale_y_reverse(name = 'Depth (cm)', limits = c(35, -1)) +
  xlab(expression(Delta^14*C[CO2] ~ ("\u2030"))) +
  geom_vline(xintercept=20.75, linetype="dashed", size=0.3) +
  geom_errorbarh(aes(xmax=X14C+X14C_sigma, xmin=X14C-X14C_sigma, colour=measurement)) +
  scale_colour_discrete(name="Profile")  
print(atmcorrplot + proftheme)
```

Plot 14C depth profiels using the non-corrected data
```{r}
profiles_noncorr <- profiles_noncorr %>% mutate(depth = ifelse(!is.na(depth), depth, 0)) %>% arrange(measurement, depth)

noncorrplot <- ggplot(profiles_noncorr, aes(x=X14C, y=depth, group=measurement)) +
  geom_point(aes(colour=measurement)) +
  geom_path(aes(colour=measurement)) +
  scale_y_reverse(name = 'Depth (cm)', limits = c(35, -1)) +
  xlab(expression(Delta^14*C[CO2] ~ ("\u2030"))) +
  geom_vline(xintercept=20.75, linetype="dashed", size=0.3) +
  geom_errorbarh(aes(xmax=X14C+X14C_sigma, xmin=X14C-X14C_sigma, colour=measurement)) +
  scale_colour_discrete(name="Profile") 
print(noncorrplot + proftheme)
ggsave("profile14C_noncorr.png", path="plots", width = 6,height = 4, units="in", dpi = 300)
```

Plot each profile individually, using non-corrected data
```{r}
for(i in 1:length(unique(profiles_noncorr$measurement))){
  id <- unique(profiles_noncorr$measurement)[i]

  profileplot <- ggplot(profiles_noncorr %>% filter(measurement==id), aes(x=X14C, y=depth)) + 
    geom_point() +
    ggtitle(id) +
    scale_y_reverse(name = 'Depth (cm)', limits = c(30, -5)) +
   xlab(expression(Delta^14*C[CO2] ~ ("\u2030"))) +
   geom_errorbarh(aes(xmax=X14C+X14C_sigma, xmin=X14C-X14C_sigma), height=2) +
   geom_path()
  print(profileplot + proftheme)
}
```